### 2022-05-19 531_stars/exam2/exam.tex
set up question

\documentclass[12pt, letterpaper, twoside]{article}
\usepackage{nopageno,epsfig, amsmath, amssymb}
\usepackage{physics}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{mhchem}
\usepackage[normalem]{ulem}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}
\usepackage{empheq}
\usepackage{wrapfig}
\usepackage[shortlabels]{enumitem}

\usepackage[letterpaper,
            margin=0.8in]{geometry}

\newcommand{\psetnum}{2}
\newcommand{\class}{ASTR 531 - Stellar Interiors and Evolution}

\newcommand{\tomtitle}{
    \noindent {\LARGE \fontfamily{cmr}\selectfont \textbf{\class}} \hfill \\[1\baselineskip]
    \noindent {\large \fontfamily{cmr}\selectfont Exam \psetnum \hfill \textsc{Tom Wagg}}\\[0.5\baselineskip]
    {\fontfamily{cmr}\selectfont \textit{\today}}\\[2\baselineskip]
}

\title{\class : Exam \psetnum}
\author{\textbf{Tom Wagg}}

\newcommand{\question}[1]{{\noindent \it #1}}
\newcommand{\answer}[1]{
    \par\noindent\rule{\textwidth}{0.4pt}#1\vspace{0.5cm}
}
\newcommand{\todo}[1]{{\color{red}\begin{center}TODO: #1\end{center}}}

% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\newcommand{\avg}[1]{\left\langle #1 \right\rangle}
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\allowdisplaybreaks

\begin{document}

\tomtitle

\vspace{-0.5cm}

\noindent \textbf{[Exam format]}

\noindent This exam is open note but closed book/internet. You have access to a list of constants and maybe use a tool for converting units (e.g.\ an iPython instance with astropy installed).

\section*{Binary Limbo: How low can you(r separation) go?!}

\question{For each type of mass transfer listed below state whether there is a minimum separation that a binary can achieve during mass transfer.\\
If you answer yes
\begin{itemize}
    \item Explain what physically occurs in the system when it reaches this minimum separation
    \item Find an expression for the minimum separation in terms of the initial separation and initial mass ratio $q = M_{2, i} / M_{1, i}$ (assuming that the rate of mass transfer is constant)
\end{itemize}
If you answer no
\begin{itemize}
    \item Explain why a minimum separation does not exist
\end{itemize}
}

\question{\textbf{Part a - Conservative and Stable Mass Transfer}}
\answer{}

\question{\textbf{Part b - Unstable Mass Transfer}}
\answer{}

\end{document}