### 2022-10-10 541_ism/week2/solutions.tex
week 2 problem 1 astr541

\documentclass[12pt, letterpaper, twoside]{article}
\usepackage{nopageno,epsfig, amsmath, amssymb}
\usepackage{physics}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{mhchem}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}
\usepackage{empheq}
\usepackage{wrapfig}

\usepackage[letterpaper,
            margin=0.8in]{geometry}

\newcommand{\psetnum}{2}
\newcommand{\class}{ASTR 541 - Interstellar Medium}

\newcommand{\tomtitle}{
    \noindent {\LARGE \fontfamily{cmr}\selectfont \textbf{\class}} \hfill \\[1\baselineskip]
    \noindent {\large \fontfamily{cmr}\selectfont Problem Set \psetnum \hfill \textsc{Tom Wagg}}\\[0.5\baselineskip]
    {\fontfamily{cmr}\selectfont \textit{\today}}\\[2\baselineskip]
}

\title{\class : Week \psetnum}
\author{\textbf{Tom Wagg}}

\newcommand{\question}[1]{{\noindent \it #1}}
\newcommand{\answer}[1]{
    \par\noindent\rule{\textwidth}{0.4pt}#1\vspace{0.5cm}
}
\newcommand{\todo}[1]{{\color{red}\begin{center}TODO: #1\end{center}}}

% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\newcommand{\avg}[1]{\left\langle #1 \right\rangle}
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\allowdisplaybreaks

\begin{document}

\tomtitle{}

\question{\textbf{1. Maxwellians}}

\answer{
    The minimum velocity needed to ionise a hydrogen atom can be calculated by setting the kinetic energy equal to the ionisation energy
    \begin{align}
        \frac{1}{2} m_{e} v_{\rm ionise}^2 &= E_{\rm ionise} \\
        v_{\rm ionise} &= \sqrt{\frac{2 E_{\rm ionise}}{m_e}} \\
        \Aboxed{ v_{\rm ionise} &= 2187 \unit{km}{s^{-1}}}
    \end{align}
    We now need to integrate the Maxwellian (yay?). Now I don't intend to do that analytically because I am not sadistic and therefore will be numerically integrating it will \texttt{scipy}'s quadrature function. Additionally, it doesn't play nicely with infinity so we're going to be sneaky and flip it around
    \begin{align}
        f_{\rm ionise} &= \int_{v_{\rm ionise}}^{\infty} 4 \pi \qty(\frac{m_e}{2 \pi k_B T})^{\frac{3}{2}} \exp\qty(-\frac{m v^2}{2 k_B T}) v^2 \dd{v} \\
        f_{\rm ionise} &= 1 - \int_0^{v_{\rm ionise}} 4 \pi \qty(\frac{m_e}{2 \pi k_B T})^{\frac{3}{2}} \exp\qty(-\frac{m v^2}{2 k_B T}) v^2 \dd{v}
    \end{align}
    Now we can evaluate this for both temperatures given
    \begin{equation}
        \boxed{ f_{\rm ionise, HII} = 6.5 \times 10^{-7},\quad f_{\rm ionise, halo} = 0.96 }
    \end{equation}
    From this we can conclude that collisional ionisation is extremely unlikely to occur in an HII region, but it will be dominant for hot halo gas.
}

\question{\textbf{2. The Dust Collector}}

\answer{}

\end{document}